---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- devtools::rmarkdown::render("README.Rmd") -->
<!-- Rscript -e "library(knitr); knit('README.Rmd')" -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# WEGE index

WEGE is an R package that allows the user to calculate the WEGE index for a particular area. Additionally it also calculates rasters of KBA criteria (A1a, A1b, A1e, and B1) Weighted endemism, the EDGE score, Evolutionary Distinctiveness and Extinction risk.

# Install

The package can currently only be installed through GitHub:

```{r gh-installation, include=TRUE, echo=TRUE, eval=FALSE}
# install.packages("remotes")
remotes::install_github("harithmorgadinho/wege_ind")
```

# Usage

A get_edge example:

```{r get_edge-example, include=TRUE, eval=TRUE}
library(WEGE) 
library(sp)
library(sf)

species <- letters[1:26]
range_list <- list()
for (i in seq_along(species)) {
  temp0 <- cbind(runif(3,1,50),runif(3,1,50))
  temp <- Polygon(rbind(temp0,temp0[1,]))
  range_list[[i]] <- Polygons(list(temp), ID = c(species[i]))
}
input <- st_as_sf(SpatialPolygons(range_list))
categories <- c('LC','NT','VU','EN','CR')
input$binomial <- species
input$category <- sample(size = nrow(input),x = categories,replace = TRUE)
input$ED <- runif(nrow(input),1,30)
temp0 <- cbind(runif(3,1,50),runif(3,1,50))
target_area <- Polygon(rbind(temp0,temp0[1,]))
target_area <- Polygons(list(target_area), ID = 'Target area')
target_area <- st_as_sf(SpatialPolygons(list(target_area)))

get_edge(target_area = target_area,input = input,species = 'binomial',category = 'category')
```

A get_wege example:

```{r get_wege-example, include=TRUE, eval=TRUE}
library(sp)
library(sf)
library(WEGE)

species <- letters[1:26]
range_list <- list()
for (i in seq_along(species)){
  temp  <-  Polygon(cbind(runif(4,1,50),runif(4,1,50)))
  range_list[[i]] <- Polygons(list(temp), ID = c(species[i]))}
input <- st_as_sf(SpatialPolygons(range_list))
categories <- c('LC','NT','VU','EN','CR')
input$binomial <- species
input$category <- sample(size = nrow(input),x = categories,replace = TRUE)

target_area <- Polygon(cbind(runif(4,1,50),runif(4,1,50)))
target_area <- Polygons(list(target_area), ID = 'Target area')
target_area <- st_as_sf(SpatialPolygons(list(target_area)))
get_wege(target_area,input,species = 'binomial',category = 'category')
```

A get_kba-criteria example:

```{r get_kba_criteria-example, include=TRUE, eval=TRUE}
library(WEGE)
library(sp)
library(sf)

species <- letters[1:26]
range_list <- list()
for (i in seq_along(species)){
  temp0 <- cbind(runif(3,1,50),runif(3,1,50))
  temp  <-  Polygon(rbind(temp0,temp0[1,]))
  range_list[[i]] <- Polygons(list(temp), ID = c(species[i]))
}
input <- st_as_sf(SpatialPolygons(range_list))
categories <- c('LC','NT','VU','EN','CR')
input$binomial <- species
input$category <- sample(size = nrow(input),x = categories,replace = TRUE)

temp0 <- cbind(runif(3,1,50),runif(3,1,50))
target_area <- Polygon(rbind(temp0,temp0[1,]))
target_area <- Polygons(list(target_area), ID = 'Target area')
target_area <- st_as_sf(SpatialPolygons(list(target_area)))
get_kba_criteria(target_area,input)

```

A raster example example:

```{r raster-example, include=TRUE, eval=TRUE}
library(WEGE)
library(sp)
library(sf)
library(raster)

 species <- sample(letters, 10)
 range_list <- list()
 for (i in seq_along(species)) {
   temp0 <- cbind(runif(3,1,50),runif(3,1,50))
   temp <- Polygon(rbind(temp0,temp0[1,]))
   range_list[[i]] <- Polygons(list(temp), ID = c(species[i]))
   }
 input <- st_as_sf(SpatialPolygons(range_list))
 categories <- c('LC','NT','VU','EN','CR')
 input$binomial <- species
 input$category <- sample(size = nrow(input),x = categories,replace = TRUE)
 
 input$ed <- runif(runif(10,1,50))
 temp0 <- cbind(runif(3,1,50),runif(3,1,50))
 target_area <- Polygon(rbind(temp0,temp0[1,]))
 target_area <- Polygons(list(target_area), ID = 'Target area')
 target_area <- st_as_sf(SpatialPolygons(list(target_area)))
 spat_ras(target_area,input,species = 'binomial',ed='ed', res = 1)
```

# Citation:
Farooq, H., Azevedo, J., Belluardo F., Nanvonamuquitxo, C., Bennett, D., Moat, J., Soares, A., Faurby, S. & Antonelli, A. (2020). Wege: A New Metric for Ranking Locations for Biodiversity Conservation. bioRxiv. https://doi.org/10.1101/2020.01.17.910299
